---
title: "Replication of David Schoch's blog post 'Six degrees of Hadley Wickham: The CRAN co-authorship network'"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article aims to replicate David Schoch's analysis in his blog post ["Six degrees of Hadley Wickham: The CRAN co-authorship network"](https://blog.schochastics.net/posts/2024-01-17_six-degrees-of-hadley-wickham/index.html) using the `authoritative` package.

```{r setup}
library(authoritative)
library(tidyverse)
library(igraph)
library(netUtils)
```

```{r}
db <- tools::CRAN_package_db()
author_pkg_cran <- db |> 
  dplyr::transmute(
    Package = Package,
    parsed_authors_r = purrr::map(authoritative::parse_authors_r(`Authors@R`), format, include = c("given", "family")),
    parsed_authors = authoritative::parse_authors(Author)
  ) |> 
  dplyr::mutate(
    authorsR = coalesce(parsed_authors_r, parsed_authors),
    authorsR = map(authorsR, stringi::stri_trans_general, "ASCII"),
    authorsR = map(authorsR, stringr::str_replace_all, "[[:punct:]]", " "),
    authorsR = map(authorsR, stringr::str_replace_all, "[[:space:]]+", " "),
    authorsR = map(authorsR, trimws),
    .keep = "unused"
  ) |> 
  tidyr::unnest_longer(authorsR) |> 
  dplyr::mutate(authorsR = authoritative::expand_names(authorsR, authorsR)) |> 
  # Standardize capitalization
  dplyr::mutate(
    authorsR = stringr::str_to_title(authorsR)
  )
```

```{r}
# FIXME: these signal an problem in the parsing. So this should be fixed instead of dropping the values
author_pkg_cran <- author_pkg_cran |> 
  filter(!authorsR %in% Package)
```

```{r}
author_pkg_cran_net <- netUtils::bipartite_from_data_frame(author_pkg_cran, "authorsR", "Package")
A <- as_biadjacency_matrix(author_pkg_cran_net, sparse = TRUE)
A <- as(A, "sparseMatrix")
B <- Matrix::t(A) %*% A
auth_auth_net <- graph_from_adjacency_matrix(B, "undirected", diag = FALSE, weighted = TRUE)
auth_auth_net
```

```{r}
comps_cran <- components(auth_auth_net)
comps_cran$no
```

```{r}
auth_auth_net_largest <- largest_component(auth_auth_net)
auth_auth_net_largest
```

> From the `r format(length(unique(author_pkg_cran$authorsR)),big.mark=",")` recorded package authors, `r format(vcount(auth_auth_net_largest),big.mark=",")` (`r round(vcount(auth_auth_net_largest)/length(unique(author_pkg_cran$authorsR))*100,2)`%) are part of the largest connected component. All subsequent analyses will be done with this network.

```{r}
emax_who <- ends(auth_auth_net_largest, E(auth_auth_net_largest)[which.max(E(auth_auth_net_largest)$weight)])[1, ]
emax <- max(E(auth_auth_net_largest)$weight)
```

> On average, every developer in the largest component has `r round(mean(degree(auth_auth_net_largest)),2)` co-authors. The median is `r median(degree(auth_auth_net_largest))`. 
The two individuals who coauthored the most packages together (`r emax`), are `r emax_who[1]` and `r emax_who[2]`. The person with the most co-authors (`r max(degree(auth_auth_net_largest))`) is `r V(auth_auth_net_largest)$name[which.max(degree(auth_auth_net_largest))]`. What a great transition for the next section.

```{r}
hadley_idx <- which(V(auth_auth_net_largest)$name == "Hadley Wickham")
author_tbl <- tibble(
    name = V(auth_auth_net_largest)$name,
    dist2HW = distances(auth_auth_net_largest, to = hadley_idx, weights = NA),
    cc = (1 / closeness(auth_auth_net_largest, weights = NA)) * 1 / vcount(auth_auth_net_largest)
)
```

```{r}
author_tbl |>
    count(dist2HW) |>
    ggplot(aes(x = dist2HW, y = n)) +
    geom_col(color = "white", fill = "black") +
    theme_minimal() +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
    ) +
    scale_x_continuous(breaks = seq(0, 10), name = "Hadley Number")
```

```{r}
author_tbl |>
    select(name, centrality = cc) |>
    dplyr::top_n(10, -centrality) |>
    dplyr::arrange(centrality) |>
    knitr::kable()
```